\chapter{Hardware Setup}
\label{ch:hardware-setup}

This chapter describes the recommended hardware setup for the
\textbf{MICSBOL/ESP32\_BT\_Controller} firmware: the ESP32 receives control packets from the Android app over
Bluetooth (SPP), drives actuators through the hardware abstraction layer (HAL), and can optionally read
sensor inputs for telemetry.

\section{ESP32 Development Board}
\label{sec:hw-esp32-board}

The ESP32 acts as the main controller. After flashing the firmware and powering the board, it advertises a
Bluetooth device named \texttt{ESP32-BT}. If Android prompts for a PIN during pairing, the default PIN is
\texttt{1234}. (Bluetooth profile: SPP.)\footnote{See the repository's \texttt{README.md} connection guide for the
    exact device name, default PIN, and quick reference.}

A typical setup includes:
\begin{itemize}
    \item an ESP32 development board flashed with the repository firmware,
    \item a stable 5V supply (USB power bank / charger or regulated 5V source),
    \item connected actuators (motor driver, servos, LEDs/buzzer), and
    \item optional sensors (analog or I\textsuperscript{2}C) for telemetry.
\end{itemize}

\paragraph{Project modes}
The firmware supports different wiring approaches depending on the selected mode in \texttt{user\_config.h}:
\begin{itemize}
    \item \texttt{MODE\_FULL\_RC} (default): uses ESP32 built-in pins for outputs.
    \item \texttt{MODE\_FULL\_RC\_MCP}: expands digital outputs using an MCP23017 GPIO expander.
\end{itemize}

\section{Control Outputs (What the App Controls)}
\label{sec:hw-control-outputs}

Once connected, the Android app provides control channels that are mapped in the firmware's control/HAL layer:
\begin{itemize}
    \item Left stick X: steering (left/right)
    \item Left stick Y: left motor speed
    \item Right stick X: camera pan (left/right)
    \item Right stick Y: right motor speed
    \item Left knob: LED brightness (0 = off, max = full)
    \item Right knob: buzzer intensity (0 = silent, max = full)
    \item Switches 1--6: digital outputs (on/off)
    \item Event buttons: send a short pulse to a switch output
\end{itemize}

These mappings are summarized in the repository \texttt{README.md} and implemented by the firmware in
\texttt{control.cpp} and \texttt{hal\_output.cpp}.

\section{Motor Driver Connection}
\label{sec:hw-motor-driver}

DC motors must \textbf{not} be powered directly from ESP32 pins. A motor driver is required (for example,
a dual H-bridge module).

The driver typically needs:
\begin{itemize}
    \item \textbf{Motor supply (VM)}: battery voltage suitable for the motors,
    \item \textbf{Logic supply (VCC)}: usually 3.3V or 5V depending on the driver module,
    \item \textbf{GND}: common ground shared with the ESP32, and
    \item \textbf{Control inputs}: PWM and direction inputs driven by ESP32 GPIOs (through the HAL).
\end{itemize}

\paragraph{Common ground is mandatory}
If the motor driver and the ESP32 do not share a common ground, PWM and direction signals will be unreliable
and motors may not respond.

\section{Servo Connections}
\label{sec:hw-servos}

Servo motors can be controlled from ESP32 PWM-capable GPIO pins. The firmware treats some channels as PWM
outputs that may be mapped to steering and camera pan.

\paragraph{Servo power}
Even though the PWM signal comes from the ESP32, most servos should be powered from a separate 5V rail:
\begin{itemize}
    \item Servo \textbf{V+}: external 5V regulator (sized for the servo stall current)
    \item Servo \textbf{GND}: shared with ESP32 GND
    \item Servo \textbf{Signal}: ESP32 PWM pin
\end{itemize}

\section{Optional Inputs and Telemetry Sensors}
\label{sec:hw-telemetry-inputs}

The firmware can send telemetry back to the Android app (numeric panels, indicators, and plots). The repository
describes telemetry elements such as panels, indicators (throttle/battery), and plots (volts/amps/RPM). Telemetry
values can come from real sensors (analog/I\textsuperscript{2}C) or can be injected for testing using the telemetry
debug modes in \texttt{debug\_config.h}.

Typical sensor interfaces:
\begin{itemize}
    \item \textbf{Analog inputs (ADC)}: for battery monitors, potentiometers, temperature sensors, etc.
    \item \textbf{I\textsuperscript{2}C}: for external sensors or I/O expansion (MCP23017 in MCP mode)
\end{itemize}

\section{Power Supply and Safety}
\label{sec:hw-power}

A reliable power system is essential for stable Bluetooth operation and correct motor/servo control.

\subsection{Recommended power arrangement}
\begin{itemize}
    \item \textbf{ESP32:} powered from 5V USB or a regulated 5V rail.
    \item \textbf{Motors:} powered from a motor battery rail through a motor driver (VM).
    \item \textbf{Servos (if used):} powered from a separate 5V regulator capable of high peak current.
\end{itemize}

\subsection{Grounding}
Always connect the grounds together:
\begin{itemize}
    \item ESP32 GND
    \item motor driver GND
    \item servo power GND
    \item sensor GND
\end{itemize}

\subsection{Practical notes}
\begin{itemize}
    \item Keep motor power wiring physically separated from ESP32 signal wiring when possible.
    \item Add decoupling capacitors near motor drivers/servo rails to reduce brownouts.
    \item If the ESP32 resets when motors start, the cause is usually power sag or noise: improve the regulator,
    add capacitance, and verify the ground wiring topology.
\end{itemize}